Open Data, Open Source and Open Standards in chemistry: \\ The Blue Obelisk five years on

### 2011-06-10 paper.tex
Added text from Andy Lang and JC Bradley

>>>>

\subsection*{Reaction Attempts}
Although there are existing databases that allow for searching reactions, those
using Open Data are harder to find. The Reaction Attempts database
\cite{ReactionAttempts}, to which anyone can
submit reaction attempts data, consists mainly of reaction information
abstracted from Open Notebooks in organic chemistry, such as the
UsefulChem project from the Bradley group
\cite{UsefulChemReactionAttempts}
and the notebooks from the Todd group \cite{UsefulChemTodd}.
Key information from each experiment is abstracted manually, with the
only required information consisting of the ChemSpider IDs of the
reactants and the product targeted in the experiment; and a link to
the laboratory notebook page. Information in the database can be
searched and accessed using the web-based Reaction Attempts Explorer
\cite{ReactionAttemptsExplorer}.

Since the database reflects all data
from the notebooks, it includes experiments in progress, ambiguous results and
failed runs. Unlike most reaction databases that only identify
experiments successfully reported in the literature, 
the Reaction
Attempts Explorer allows researchers to easily find patterns in
reactions that have already been performed, 
and since the data are open
and results are reported across all research groups, intersections are
easily discovered and possible Open Collaboration opportunities are
easily found \cite{UsefulChemSocialNetworks, ekins_collaboration_2011}.

\subsection*{Non-Aqueous Solubility}

Although the aqueous solubility of many common organic compounds is
generally available, quantitative reports of non-aqueous solubility
are more difficult to find.
Such information can be valuable for selecting solvents for
reactions, re-crystallization and related processes. In 2008, the Open
Notebook Science Solubility Challenge was launched for the purpose of
measuring non-aqueous solubility of organic compounds, reporting all
the details of the experiments in an Open Notebook and recording the
results as Open Data in a centralized
database~\cite{ONS2010, BeautifulData_2009}.
This crowdsourcing project was also supported by Submeta,
Sigma-Aldrich, Nature Publishing Group and the Royal Society of
Chemistry. The database currently holds 1932 total measurements and
1428 averaged solute/solvent measurements all of which are available under
a CC0 license.  Several web services and
feeds are available to filter and re-use the dataset~\cite{SolubilityServices}.
In particular, models have been developed for the prediction of
non-aqueous solubility in 72 different
solvents~\cite{UCSolubilityPrediction}
using the method of Abraham et al~\cite{AbrahamSolubility} with
descriptors calculated by
the Chemistry Development Kit. These models are available online and
will be refined as more solubility data is collected.
